Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-01-29 apj.tex
Painful rewriting of a section

@@

 \author{    D.~Rubin}
 \affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
    \affiliation{    Department of Physics, Florida State University,
    315 Keen Building, Tallahassee, FL 32306-4350}

@@

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter is responsible for supernova color diversity
beyond those tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset, $\Delta$, is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  
The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
Unlike the parameters associated with
the spectral measurements, $EW_{Ca}$, $EW_{Si}$ and $\lambda_{Si}$,  the latent
parameters $k_0$ and $k_1$ are not directly associated
with observables but are rather drawn from the set of peak magnitudes.
The model 
\color{red}
likelihood density
\color{black}
is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} \\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si}  \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} \\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} \\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{0} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ and $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

To constrain the degrees of freedom and degeneracies inherent in the model we impose that
each of the vectors $k_0+1/N$ and $k_1+1/N$ are simplexes (all elements are $\ge 0$ and sum to one),
where $N$ is the number of objects in our sample: these
conditions break the 
degeneracies  $\gamma ^0\rightarrow a\gamma^0$, $k_0 \rightarrow a^{-1}k_0$ and $\gamma^1 \rightarrow a\gamma^1$, $k_1 \rightarrow a^{-1}k_1$
up to a $\pm$ sign.
Furthermore we impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0, \gamma^0_0 > 0, \gamma^1_0 < 0.
\end{equation}
The first two conditions specify the definition of zero color relative to which the color excess is measured:
in a linear model this value is arbitrary.
The latter two exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_0$ and $\gamma^1_0$ distinguishes the two.)
As  $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other,
the different initial and boundary conditions for $\gamma^0_0$ and $\gamma^1_0$ break the degeneracy of our mixture model;
otherwise identical initial conditions for $\gamma^0$, $\gamma^1$ produce indistinguishable posteriors
with relatively broad credible intervals and lower maximum likelihood.

@@

The range of SiII$\lambda$4130 equivalent widths is $\pm 20$~\AA\ whereas the width of the $B$-band is 851~\AA, so that its affect on magnitude
is
$2.5 \log{(20/850)} \sim 0.03$ mag.  
The implied span in $B$ magnitude based on $\beta_1$ is 0.54~mag.  Therefore $\beta_1$ cannot wholly be attributed to the flux deficit
from the line itself.
Similarly, the CaII H\&K equivalent widths have range $\pm 50$~\AA, while the width of the $U$ band is
701~\AA, so that its affect on magnitude
is
$2.5 \log{(50/701)} \sim 0.08$ mags.   The implied span in $U$ magnitude is  0.21~mag, so $\alpha_0$ cannot be completely due to the line itself.

@@

To visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that the four vectors point in different directions, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.

The specific directions of $\frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}$ and
$\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}$ within the plane they define are not arbitrary, but are determined by the prior that
$k_0$ and  $k_1$ are uncorrelated. 
In order to understand these directions in the context of the dust model,
we
rewrite the \citetalias{1999PASP..111...63F} model in Eqn.~\ref{f99:eqn} as
\begin{equation}
A_X =  A_V^F\left(a(X)+\frac{b(X)}{r} \right) +  \left( E^F(B-V) - \frac{ A_V^F}{r}\right)b(X).
\end{equation}
Assigning $r=2.4$, we find that the combination  $\left(a(X) + \frac{b(X)}{2.4}\right)$ is
almost perfectly aligned with $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$ while $b$ alone is almost perfectly aligned with $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$,
as shown in the dot-dashed and dotted lines of  Figure~\ref{plane:fig}.  This implies a correspondence of $k_0$ with $A^F_V$,
and $k_1$ with $  \left( E^F(B-V) - \frac{ A_V^F}{2.4}\right)$.
Our prior of a non-correlation between $k_0$ and $k_1$ has lead to an $r \sim 2.4$  that optimizes the
non-correlation between $A^F_V$ and $  \left( E^F(B-V) - \frac{ A_V^F}{r}\right)$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/plane0.pdf}
   \includegraphics[width=2.8in]{output11/plane1.pdf}
   \caption{
   Two different perspectives in $UVI$-space of the vectors $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.4$ is shown in the dot-dashed
   line, almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$, while $b(X)$ shown in the dotted line is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$.  
   \label{plane:fig}}
\end{figure}

Given the association of  our parameter $\gamma^0_V k_0$ with the $A^F_V$ of an $R^F \sim 2.4$ dust model ,
the former's distribution can be compared to the expectation for the latter.
Figure~\ref{k0_med:fig} shows the histogram of
median values of $\gamma^0_V k_0$.
The distribution is non-Gaussian, having a sharp rise in the blue and an extended tail in the red, consistent
with the expectation of dust considering the spatial distribution of supernovae within galaxies and the distribution of galaxy orientations with respect to the observer \citep{1998ApJ...502..177H}.

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which is associated with $\left( E^F(B-V) - \frac{ A_V^F}{2.4}\right)$, i.e.\
extinction corrections due to deviations away from the $R=2.4$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
The implication is that while the bulk of supernovae have positive values associated with $R>2.4$, there is a tail
of negative values associated with $R<2.4$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$. On the other hand, lower values of $\gamma^1_V k_1$
occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
\citet{2011ApJ...729...55F} and the above authors
find that supernovae with large extinction 
preferentially exhibit low $R^F$.

@@

The near coplanarity of the plane defined by our parameters and the plane of  \citetalias{1999PASP..111...63F}, together with the fact that our  supernova
parameters translate to  \citetalias{1999PASP..111...63F} parameters with consistent distributions, lead us to conclude that 
$\{k_0, k_1\}$, describe for the most part the $\{ E(B-V), R\}$ parameters used
in standard models of Milky Way dust.
\color{black}

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} + \delta_0 D\\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si} + \delta_1 D \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} + \delta_2 D\\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} + \delta_3 D\\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si} + \delta_4 D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{0} k_0 +\gamma^1_{0} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes:
each supernova has an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

>>>>

\author{ S.~Dixon}
\affiliation{
    Department of Physics, University of California Berkeley,
    366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}

@@

 \author{    D.~Rubin}
 \affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
    \affiliation{   Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218}

@@

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$1-5\times10^{-5}$
%------
confidence.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$1-5\times10^{-4}$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset, $\Delta$, is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  
The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
Unlike the parameters associated with
the spectral measurements, $EW_{Ca}$, $EW_{Si}$ and $\lambda_{Si}$,  the latent
parameters $k_0$ and $k_1$ are not directly associated
with observables but are rather drawn from the set of peak magnitudes.
The model 
\color{red}
likelihood density
\color{black}
is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} \\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si}  \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} \\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} \\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{U} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ and $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

To constrain the degrees of freedom and degeneracies inherent in the model we impose that
each of the vectors $k_0+1/N$ and $k_1+1/N$ are simplexes (all elements are $\ge 0$ and sum to one),
where $N$ is the number of objects in our sample: these
conditions break the 
degeneracies  $\gamma ^0\rightarrow a\gamma^0$, $k_0 \rightarrow a^{-1}k_0$ and $\gamma^1 \rightarrow a\gamma^1$, $k_1 \rightarrow a^{-1}k_1$
up to a $\pm$ sign.
Furthermore we impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0, \gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
The first two conditions specify the definition of zero color relative to which the color excess is measured:
in a linear model this value is arbitrary.
The latter two exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)
As  $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other,
the different initial and boundary conditions for $\gamma^0_U$ and $\gamma^1_U$ break the degeneracy of our mixture model;
otherwise identical initial conditions for $\gamma^0$, $\gamma^1$ produce indistinguishable posteriors
with relatively broad credible intervals and lower maximum likelihood.

@@

The range of SiII$\lambda$4130 equivalent widths is $\pm 20$~\AA\ whereas the width of the $B$-band is 851~\AA, so that its affect on magnitude
is
$2.5 \log{(20/850)} \sim 0.03$ mag.  
The implied span in $B$ magnitude based on $\beta_B$ is 0.54~mag.  Therefore $\beta_B$ cannot wholly be attributed to the flux deficit
from the line itself.
Similarly, the CaII H\&K equivalent widths have range $\pm 50$~\AA, while the width of the $U$ band is
701~\AA, so that its affect on magnitude
is
$2.5 \log{(50/701)} \sim 0.08$ mags.   The implied span in $U$ magnitude is  0.21~mag, so $\alpha_U$ cannot be completely due to the line itself.

@@

To visualize the above result, Figure~\ref{plane:fig} shows two perspectives (left and right columns) of $UVI$- (top row) and $BVR$- (bottom row) spaces
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  
The two perspectives show that while the four vectors point in different directions for each band combination,
they are almost coplanar in $UVI$, and close to coplanar in $BVR$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11/plane0.pdf}
   \includegraphics[width=2.95in]{output11/plane1.pdf}
   \includegraphics[width=2.95in]{output11/plane0BVR.pdf}
   \includegraphics[width=2.95in]{output11/plane1BVR.pdf}
   \caption{
   Two perspectives (left and right columns) of $UVI$- (top row) and $BVR$- (bottom row) spaces
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combinations $a(X)+b(X)/2.4$ and  $a(X)+b(X)/2.6$ are shown in the dotted
   lines: thy are almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$ for $UVI$ and $BVR$ respectively.  $b(X)$ shown in the dotted line is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$ in $UVI$ but is clearly distinct in $BVR$.  
   \label{plane:fig}}
\end{figure}

A prior that $k_0$ and  $k_1$ be uncorrelated was enforced to break degeneracies in our linear models, and led to the specification of the directions of
the vectors $\gamma^0$ and $\gamma^1$.  Suppose that that the data were described by a dust model.  A \citetalias{1999PASP..111...63F}  dust model 
can be written as 
%To see how this prior would operate on dust models, consider that themodel in Eqn.~\ref{f99:eqn} can
%be expressed in terms of new basis vectors parametrized by $\kappa$:
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a(X)+\kappa_1 b(X) \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a(X) + b(X))],
\end{equation}
with the parameters $\kappa_1$ and $\kappa_2$ providing the freedom of choice in the directions of the two basis vectors.  In the analysis,
the arbitrarity is broken and the $\kappa$'s specified, by the  prior of  
zero covariance between $(A^F_V - \kappa_2 E^F(B-V))$ and  $(-\kappa_1 A^F_V + E^F(B-V))$.
The preferred values for $\kappa$ from the linear model in \S\ref{model:sec} are shown qualitatively in
Figure~\ref{plane:fig}.  For $\kappa_1=1/2.4$, $\kappa_2=-6.8$
the vector $\gamma^0_X$ is  aligned with 
the dust vector $\left(a(X)+\kappa_1 b(X) \right)$  and $\gamma^1_X$
is aligned with $(\kappa_2 a(X) + b(X))$ for $X=UVI$.  Slightly shifted values of $\kappa_1=1/2.6$, $\kappa_2=-6.1$ apply for  $X=BVR$.

An intuitive understanding of $\kappa_1$ can be obtained by considering what would happen if
all supernovae were to experience dust with the same $R^F$.   The zero-covariance
prior would be satisfied by  $\kappa_1=1/R^F$ and any arbitrary $\kappa_2$.
The combination  $\gamma^0_V k_0$ would correspond to  $A^F_V$.  Intrinsically $k_1=0$  and $\gamma^1$ would be unconstrained.
A slight relaxation away from the constant $R^F$ assumption would result in non-zero $k_1$ and a $\gamma^1$ in a direction that satisfies
the prior condition, as parameterized by $\kappa_2$.

Given the association of  our parameter $\gamma^0_V k_0$ with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model ,
the former's distribution can be compared to the expectation for the latter.
Figure~\ref{k0_med:fig} shows the histogram of
median values of $\gamma^0_V k_0$.
The distribution is non-Gaussian, having a sharp rise in the blue and an extended tail in the red, consistent
with the expectation of dust considering the spatial distribution of supernovae within galaxies and the distribution of galaxy orientations with respect to the observer \citep{1998ApJ...502..177H}.

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which is associated with 
extinction corrections due to deviations away from the canonical $R^F$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$. On the other hand, lower values of $\gamma^1_V k_1$
occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
\citet{2011ApJ...729...55F} and the above authors
find that supernovae with large extinction 
preferentially exhibit low $R^F$.

@@

The near coplanarity of the plane defined by our parameters and the plane of  \citetalias{1999PASP..111...63F}, together with the fact that our  supernova
parameters translate to  \citetalias{1999PASP..111...63F} parameters with consistent distributions, lead us to conclude that 
$\{k_0, k_1\}$, describe for the most part the $\{ E(B-V), R\}$ parameters used
in standard models of Milky Way dust.  There do remain some small inconsistencies between our numbers and expectations from dust models.
\color{black}

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} + \delta_0 D\\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si} + \delta_1 D \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} + \delta_2 D\\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} + \delta_3 D\\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si} + \delta_4 D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{B} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes:
each supernova has an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.
